|

### 2019-09-03 paper.Rmd
Thickness rewritten

The distance vector can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$. Notice that in 3d, where for each plane we have a single orthogonal direction, this defines a flat slice of hight $2*h$, matching the common conception of a slice. When going beyond 3d, the codimension of the plane is $>1$ and using the euclidean distance results in rotation invariant slicing in the orthogonal space. As a consequence the "slices" are now spherical in the orthogonal subspace, with radius $h$.

@@

Choosing the slice thickness is a compromise between what feature size can be revealed and how many points fall inside a thin slice. As dimensionality $p$ increases, the relative number of points that are inside a slice of fixed thickness $h$ will decrease. The exact relation depends heavily on the distribution of the data points. Here we present expressions for special cases to get an idea of this scaling. For practical purposes we specify a volume $\epsilon$ and then use as a cutoff $h = \epsilon^{1/(p-2)}$. XXX not yet implemented that way, based on thickness estimates.

We consider two types of rotation invariant distributions for our estimates: a hypersphere in $p$ dimensions and a $p$ dimensional normal distribution without correlation.

Let us consider points uniformly distributed in a $p$-dimensional hypersphere. The fraction of points inside a slice of thickness $2*h$ can therefore be estimated as the relative volume of the slice compared to the volume of the full hypersphere,
\begin{equation}
V_{rel} = \frac{1}{2} \frac{h^{p-2}}{r^p} (p r^2 - (p-2) h^2) \approx \frac{1}{2} (\frac{h}{R})^{p-2},
\end{equation}
where the approximation is valid when $h \ll r$, as is typically expected to be the case. Thus if we want to keep the relative number of points (i.e. $V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$. XXXX change code to this? currently $\epsilon^{1/p}$ is implemented...

XXXX need to redo integration of multivariate normal...
Considering now slicing through a multivariate normal, we can estimate the number of points in a slice by integrating the probability distribution with boundaries $\pm h$ in $p-2$. In this case the relative density scales with the error function as
\begin{equation}
\rho_{rel} = \pi^{(2-p)/2} \mathrm{erf}((h/\sqrt 2 \sigma)) ^{p-2}.
\end{equation}

XXX two plots: vrel/rhorel scaling with h for n for p = 3, 4, 5, 10, and h scaling with p for vrel = 0.001, 0.01, 0.05 and 0.1

>>>>

The distance vector can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$. Notice that in 3d, where for each plane we have a single orthogonal direction, this defines a flat slice of hight $2h$, matching the common conception of a slice. When going beyond 3d, the codimension of the plane is $>1$ and using the euclidean distance results in rotation invariant slicing in the orthogonal space. As a consequence the "slices" are now spherical in the orthogonal subspace, with radius $h$.

@@

Choosing the slice thickness is a compromise between what feature size can be revealed and how many points fall inside a thin slice. As dimensionality $p$ increases, the relative number of points that are inside a slice of fixed thickness $h$ will decrease. The exact relation depends on the distribution of the data points. Here we estimate how it scales with $h$ and $p$ for the example of points uniformly distributed in a hypersphere. The example was selected since it represents a rotation invariant uniform distribution in $p$ space. As we are interested in hollowness this type of distribution is more relevant than a multivariate normal distribution.

The fraction of points inside a slice of thickness $h$ can be estimated as the relative volume of the slice compared to the volume of the full hypersphere,
\begin{equation}
V_{rel} = \frac{1}{2} \frac{h^{p-2}}{r^p} (p r^2 - (p-2) h^2) \approx \frac{1}{2} (\frac{h}{R})^{p-2},
\end{equation}
where the approximation is valid when $h \ll r$, as is typically expected to be the case. Thus if we want to keep the relative number of points (i.e. $V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$.
